---
title: "ACN paper script pilot study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Old data

load libraries, set working directory, load ta object of pilot study
```{r}
print(date())
#library(phyloseq)
library(vegan)
library(tidyverse)
library(ggrepel)
library(ggpubr)
devtools::install_github("SWittouck/tidyamplicons", ref = "v0.2.1")
library(tidyamplicons)
setwd("C:/Users/EOerlemans/Desktop/Acne paper")


load("Data/run_tidy _olddata.Robject")
my_comparisons=list(c(1,2),c(2,3),c(3,4), c(1,4))
```

## Setting up tidy amplicons object

rename ta object, change column name 
Also change to new taxonomy, (see other functions )

```{r}
run <- run %>% reclassify_family("Lactobacillaceae", fin_refdb, sequence_var = "taxon")

v1v2r4 <-run
names(v1v2r4$samples)[names(v1v2r4$samples) == "repeat"] <- "rep"

v1v2r4$samples<-unite(v1v2r4$samples,Samplename,type,condition,participant,rep,sep="-")

```

```{r}
report_numbers(v1v2r4)
```



Add metadata but first remove 'old' metadata:

```{r}

v1v2r4$samples <- select(v1v2r4$samples,sample, Samplename)
report_numbers(v1v2r4)
tabel = read_delim("Data/metadataV1V2run4.csv", delim =";")
v1v2r4$samples <- left_join(v1v2r4$samples,tabel)
v1v2r4<-update_tidyamplicons(v1v2r4)
```

Add library sizes etc.
```{r}
v1v2r4<- add_lib_size(v1v2r4)
#NA's for samples with 0 reads
v1v2r4<-add_rel_abundance(v1v2r4)
v1v2r4<-add_taxon_name(v1v2r4)
# most abundant gets nr 1
v1v2r4<-add_pcoa(v1v2r4)
```

First look at sample composition. NA's are controls
```{r}
get_bar_plot(v1v2r4,x=sample_id)+facet_wrap(~subgroup,scales = "free_x")
```
#Working on genus level:

```{r}
v1v2r4g<-v1v2r4
v1v2r4g$taxa<-select(v1v2r4$taxa, kingdom,phylum, class, order, family, genus, taxon_id)
v1v2r4g<-merge_taxa(v1v2r4g)
```


#  Skin analysis:
Make new object with just skin samples and check if tech repeats differ:
```{r}
skn<-v1v2r4g
skn$samples<- filter(skn$samples, str_detect(group,"ACN"))
skn<- process_sample_selection(skn)
report_numbers(skn)
get_bar_plot(skn, x=sample_id)+facet_wrap(~subgroup, scales="free_x")

study<-v1v2r4
study$samples<- filter(study$samples, str_detect(study,"Y"))
study<- process_sample_selection(study)
report_numbers(study)
get_bar_plot(study)+facet_wrap(~participant, scales="free_x")

studyg<-v1v2r4g
studyg$samples<- filter(studyg$samples, str_detect(study,"Y"))
studyg<- process_sample_selection(studyg)
report_numbers(studyg)
get_bar_plot(studyg)+facet_wrap(~participant, scales="free_x")

baseline<-v1v2r4
baseline$samples<-filter(baseline$samples, str_detect(visit,"1"))
baseline<-process_sample_selection(baseline)
get_bar_plot(baseline)+facet_wrap(~subgroup, scales="free_x")

baselineg<-skn
baselineg$samples<-filter(baselineg$samples, str_detect(visit,"1"))
baselineg<-process_sample_selection(baselineg)
get_bar_plot(baselineg)+facet_wrap(~subgroup, scales="free_x")

report_numbers(baseline)
```
Generally looks ok! --> Merge samples

```{r}
v1v2r4m<-v1v2r4
v1v2r4m$sample<-select(v1v2r4m$samples, sample_id,study,group,subgroup,number,participant,visit,staal, score,pandp,comedones)
v1v2r4m<-merge_samples(v1v2r4m)
report_numbers(v1v2r4m)

v1v2r4mg<-v1v2r4g
v1v2r4mg$sample<-select(v1v2r4mg$samples, sample_id,study,group,subgroup,number,participant,visit,staal, score,pandp,comedones)
v1v2r4mg<-merge_samples(v1v2r4mg)
report_numbers(v1v2r4mg)

```

Make objects m= merged, g= genus level 
skn is all samples
study is subsetting for the pilot study
baseline is only samples of the first visit and the controls (no intervention- healty participants)
Con - only control participants (healthy)
```{r}
sknmg<-skn
sknmg$samples<-select(sknmg$samples, sample_id,acne, study,group,subgroup,participant,visit,staal, score,pandp,comedones,-techrep)
sknmg<-merge_samples(sknmg)
report_numbers(sknmg)
sknmg<- add_lib_size(sknmg)
sknmg<-add_rel_abundance(sknmg)
sknmg<-add_pcoa(sknmg)

studym<-study
studym$samples<-select(studym$samples, sample_id,acne, study,group,subgroup,participant,visit,staal, score,pandp,comedones)
studym<-merge_samples(studym)
report_numbers(studym)
studym<-add_lib_size(studym)
studym<-add_rel_abundance(studym)
studym<-add_pcoa(studym)

studymg<-studyg
studymg$samples<-select(studymg$samples, sample_id, acne, study,group,subgroup,participant,visit,staal, score,pandp,comedones)
studymg<-merge_samples(studymg)
report_numbers(studymg)
studymg<-add_lib_size(studymg)
studymg<-add_rel_abundance(studymg)
studymg<-add_pcoa(studymg)

baselinemg<-baselineg
baselinemg$samples<-select(baselinemg$samples,acne, sample_id,study,group,subgroup,participant,visit,staal, score,pandp,comedones)
baselinemg<-merge_samples(baselinemg)
report_numbers(baselinemg)
baselinemg<-add_lib_size(baselinemg)
baselinemg<-add_diversity_measures(baselinemg)
get_bar_plot(baselinemg)+facet_wrap(~subgroup, scales="free_x")

baselinem<-baseline
baselinem$samples<-select(baselinem$samples,acne, sample_id,study,group,subgroup,participant,visit,staal, score,pandp,comedones)
baselinem<-merge_samples(baselinem)
report_numbers(baselinem)
baselinem<-add_lib_size(baselinem)
baselinem<-add_rel_abundance(baselinem)
baselinem<-add_diversity_measures(baselinem)

baselineCon<-baselinemg
baselineCon$samples<-filter(baselineCon$samples,str_detect(subgroup,"CON"))
baselineCon<-process_sample_selection(baselineCon)

baselineCon2<-baselinem
baselineCon2$samples<-filter(baselineCon2$samples,str_detect(subgroup,"CON"))
baselineCon2<-process_sample_selection(baselineCon2)

```



Genera separate:
make objects
```{r}
studyLacto<-studym
studyLacto$taxa<-filter(studyLacto$taxa,str_detect(genus,"Lactobacillus"))
studyLacto<-process_taxon_selection(studyLacto)

studynoLacto<-studym
studynoLacto$taxa<-filter(studynoLacto$taxa,!str_detect(genus,"Lactobacillus"))
studynoLacto<-process_taxon_selection(studynoLacto)
studynoLacto<-add_rel_abundance(studynoLacto)

studyStaphnoLacto<-studynoLacto
studyStaphnoLacto$taxa<-filter(studyStaphnoLacto$taxa,str_detect(genus,"Staphylococcus"))
studyStaphnoLacto<-process_taxon_selection(studyStaphnoLacto)

studynoLactog<-studymg
studynoLactog$taxa<-filter(studynoLactog$taxa,!str_detect(genus,"Lactobacillus"))
studynoLactog<-process_taxon_selection(studynoLactog)

baselineLacto<-baselinem
baselineLacto$taxa<-filter(baselineLacto$taxa,str_detect(genus,"Lactobacillus"))
baselineLacto<-process_taxon_selection(baselineLacto)



studyStaph<-studym
studyStaph$taxa<-filter(studyStaph$taxa,str_detect(genus,"Staphylococcus"))
studyStaph<-process_taxon_selection(studyStaph)

baselineStaph<-baselinem
baselineStaph$taxa<-filter(baselineStaph$taxa,str_detect(genus,"Staphylococcus"))
baselineStaph<-process_taxon_selection(baselineStaph)

studyCuti<-studym
studyCuti$taxa<-filter(studyCuti$taxa,str_detect(genus,"Propionibacterium"))
studyCuti<-process_taxon_selection(studyCuti)
studyCuti<-add_taxon_name(studyCuti)

studyCutinoLacto<-studynoLacto
studyCutinoLacto$taxa<-filter(studyCutinoLacto$taxa,str_detect(genus,"Propionibacterium"))
studyCutinoLacto<-process_taxon_selection(studyCutinoLacto)

baselineCuti<-baselinem
baselineCuti$taxa<-filter(baselineCuti$taxa,str_detect(genus,"Propionibacterium"))
baselineCuti<-process_taxon_selection(baselineCuti)

```

add gender classifier 
```{r}
baselineCon$samples<-mutate(baselineCon$samples,gender=participant)
baselineCon$samples$gender<-c("F","F","F","F","F","F","F","F","F","F","F","F","F","F","F","M","M","M","M","M","M","M","M","M","M","M","M","M","M")
baselineCon2$samples<-mutate(baselineCon2$samples,gender=participant)
baselineCon2$samples$gender<-c("F","F","F","F","F","F","F","F","F","F","F","F","F","F","F","M","M","M","M","M","M","M","M","M","M","M","M","M","M")
baselineCon<-add_presence_counts(baselineCon,gender)

baselineConLacto<-baselineCon2
baselineConLacto$taxa<-filter(baselineConLacto$taxa,str_detect(genus,"Lactobacillus"))
baselineConLacto<-process_taxon_selection(baselineConLacto)

write.csv2(get_abundances_extended(baselineCon),"baselineCon_genuslevel.csv")
write.csv2(get_abundances_extended(baselineCon2),"baselineCon_ASVlevel.csv")
write.csv2(get_abundances_extended(studymg),"studymg.csv")
```

check the taxonomy through barplots for lactobacilli 
```{r}
get_bar_plot(studyLacto,x=visit)+
  facet_wrap(~participant)+
  labs(title="Lactobacillus ASVs",x="Visit",y="Relative abundance")+
  theme_minimal()+theme(legend.position = "right")
ggsave("StudyLacto_ASVcolor.png")

get_bar_plot(baselineLacto,x=as.factor(participant))+
  facet_wrap(~acne,scales = "free_x")+
  labs(title="Lactobacillus ASVs",x="Acne sample",y="Relative abundance")+
  theme_minimal()+theme(legend.position = "bottom",axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14))
ggsave("baselineLacto.png")
```

Do the same for only the controls, for staphylococci and cutibacteria (also Propionibacterium)
```{r}
get_bar_plot(baselineConLacto,x=as.factor(participant))+
  facet_wrap(~gender,scales = "free_x")+
  labs(title="Lactobacillus ASVs",x="Gender",y="Relative abundance")+
  theme_minimal()+theme(legend.position = "bottom",axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14))
ggsave("baselineLactoCon.png")

get_bar_plot(studyStaph,x=visit)+
  facet_wrap(~participant)+
  labs(title="Staphylococcus ASVs",x="Visit",y="Relative abundance")+
  theme_pubclean()+theme(legend.position = "right")
ggsave("studyStaph.png")

get_bar_plot(studyStaphnoLacto,x=visit)+
  facet_wrap(~participant)+
  labs(title="Staphylococcus ASVs (filtered)",x="Visit",y="Relative abundance")+
  theme_pubclean()+theme(legend.position = "right")
ggsave("studyStaph_LactoFiltered.png")

get_bar_plot(baselineStaph,x=as.factor(participant))+
  facet_wrap(~acne,scales = "free_x")+
  labs(title="Staphylococcus ASVs",x="Acne sample",y="Relative abundance")+
  theme_pubclean()+theme(legend.position = "bottom",axis.text.x= element_text(size = 8,colour="black",angle=90))
ggsave("baselineStaph.png")

get_bar_plot(studyCuti,x=visit)+
  labs(title="Propionibacterium and Cutibacterium ASVs",x="Visit",y="Relative abundance")+
  facet_wrap(~participant)+theme_pubclean()+theme(legend.position = "right")
ggsave("studyCuti.png")

get_bar_plot(studyCutinoLacto,x=visit)+
  labs(title="Propionibacterium and Cutibacterium ASVs (filtered)",x="Visit",y="Relative abundance")+
  facet_wrap(~participant)+theme_pubclean()+theme(legend.position = "right")
ggsave("studyCuti_LactoFiltered.png")

get_bar_plot(baselineCuti,x=as.factor(participant))+
  facet_wrap(~acne,scales = "free_x")+
  labs(title="Cutibacterium ASVs",x="Acne sample",y="Relative abundance")+
  theme_pubclean()+theme(legend.position = "bottom",axis.text.x= element_text(size = 8,colour="black",angle=90))
ggsave("baselineCuti.png")
```

Merge the ASVs that are derivate of the same applied lactobacillus
```{r}
studyLacto<-add_taxon_name_color(studyLacto)
studyLacto$taxa<-mutate(studyLacto$taxa,taxon_name_color2=taxon_name_color)
studyLacto$taxa$taxon_name_color2<-replace(studyLacto$taxa$taxon_name_color2,studyLacto$taxa$taxon_name_color2=="Lactobacillus 2","Lactobacillus 2")
studyLacto$taxa$taxon_name_color2<-replace(studyLacto$taxa$taxon_name_color2,studyLacto$taxa$taxon_name_color2=="Lactobacillus 4","Lactobacillus 2")
studyLacto$taxa$taxon_name_color2<-replace(studyLacto$taxa$taxon_name_color2,studyLacto$taxa$taxon_name_color2=="Lactobacillus 3","Lactobacillus 3")
studyLacto$taxa$taxon_name_color2<-replace(studyLacto$taxa$taxon_name_color2,studyLacto$taxa$taxon_name_color2=="Lactobacillus 5","Lactobacillus 3")
```


Make barplots for figure 4 

```{r, fig.height=3,fig.width=10}
get_bar_plot(studyLacto,geom_bar = F)+
  geom_bar(stat = "identity", aes(x= visit,fill=taxon_name_color2))+
  facet_wrap(~participant)+
  labs(title="Lactobacillus ASVs",x="Visit",y="Relative abundance")+
  theme_minimal()+theme(legend.position = "right")
ggsave("studyLacto_SpeciesColoured.png")

Lacbar<-get_bar_plot(studyLacto,geom_bar = F)+
  geom_bar(stat = "identity", aes(x= visit,fill=taxon_name_color2))+
  facet_wrap(~participant,nrow=1)+
  labs(title="Lactobacillus ASVs",x="Visit",y="Relative abundance")+
  theme_minimal()+theme(legend.position = "right")+theme(axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14), legend.text = element_text(size=11))+
  scale_y_reverse()
ggsave("studyLacto_speciesColoured_bottomLedgend.svg")
Lacbar
```

```{r}
get_bar_plot(baselineCon,x=as.factor(participant))+
  facet_wrap(~gender,scales = "free_x")+
  labs(title="Lactobacillus ASVs",x="Gender",y="Relative abundance")+
  theme_pubclean()+theme(legend.position = "bottom",axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14))
ggsave("baselineLactoCon.png")
```



Try to visualise relative abundances over visits of different samples (1 dot) for lactobacilli, staphylococci and propionibacterium/cutibacterium

```{r}
studymg$samples$visit<-as.factor(studymg$samples$visit)
studymg$samples$participant<-as.factor(studymg$samples$participant)
studymg<-add_presence_counts(studymg,visit)

relabunLac<-get_abundances_extended(studymg)
relabunLac<-filter(relabunLac,str_detect(genus,"Lactobacillus"))

baselinemg<-add_rel_abundance(baselinemg)
baselinemg<-add_presence_counts(baselinemg,acne)
relabunLacBL<-get_abundances_extended(baselinemg)
relabunLacBL<-filter(relabunLacBL,str_detect(genus, "Lactobacillus"))

baselineCon<-add_rel_abundance(baselineCon)
relabunLacCon<-get_abundances_extended(baselineCon)
relabunLacCon<-filter(relabunLacCon,str_detect(genus, "Lactobacillus"))

relabunStaph<-get_abundances_extended(studymg)
relabunStaph<-filter(relabunStaph,str_detect(genus,"Staphylococcus"))

relabunStaphBL<-get_abundances_extended(baselinemg)
relabunStaphBL<-filter(relabunStaphBL,str_detect(genus, "Staphylococcus"))

relabunCuti<-get_abundances_extended(studymg)
relabunCuti<-filter(relabunCuti,str_detect(genus,"Propionibacterium|Cutibacterium"))


relabunCutiBL<-get_abundances_extended(baselinemg)
relabunCutiBL<-filter(relabunCutiBL,str_detect(genus, "Propionibacterium"))

```

# Evolution of relative abundances
Figures with relative abundances of important genera (e.g. Figure4B)
```{r}

ggplot(relabunLac,aes(x=as.factor(visit),y=rel_abundance))+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=0.7)+
  stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.7)+
  stat_compare_means(comparisons = my_comparisons)+
  labs(x="Visit", y="relative abundance", title= "Lactobacillus")+
  theme_pubclean()+scale_y_log10()
ggsave("RelativeAbundancesLactobacilli_study.png")

BarStaphstudy<-ggplot(relabunStaph,aes(x=visit,y=rel_abundance))+
  geom_boxplot(alpha=0.5,fill='lightgrey',outlier.size = 0)+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=1,alpha=0.7,aes(col= subgroup, fill=subgroup))+
  scale_color_manual(values=c("#00798c"))+
  scale_fill_manual(values=c("#00798c"))+
  labs(x="Visit", y="relative abundance", title= "Staphylococcus-study")+
  theme_pubclean()+scale_y_log10()
ggsave("RelativeAbundancesStaphylococci_study.png")

BarCutistudy<-ggplot(relabunCuti,aes(x=visit,y=rel_abundance))+
  geom_boxplot(alpha=0.5,fill='lightgrey',outlier.size = 0)+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=1,alpha=0.7,aes(col= subgroup, fill=subgroup))+
  scale_color_manual(values=c("#00798c"))+
  scale_fill_manual(values=c("#00798c"))+
  labs(x="Visit", y="relative abundance", title= "Cutibacterium-study")+
  theme_pubclean()+scale_y_log10()+theme(legend.position = "bottom")
ggsave("RelativeAbundancesCutibacterium_study.png")

ggplot(relabunLacBL,aes(x=acne,y=rel_abundance))+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=0.7)+
  stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.7)+
  stat_compare_means(label.x.npc="middle",label.y.npc="top")+
  labs(x="group", y="relative abundance", title= "Lactobacillus")+
  theme_pubclean()+scale_y_log10(breaks = c(0.0001,0.001,0.01,0.1,1))
ggsave("RelativeAbundancesLactobacilli_baseline.png")

ggplot(relabunLacCon,aes(x=gender,y=rel_abundance))+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=0.7)+
  stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.7)+
  scale_color_manual(values=c("#66a182","#2e4057","#00798c"))+
  scale_fill_manual(values=c("#66a182","#2e4057","#00798c"))+
  stat_compare_means(label.x.npc="middle",label.y.npc="top")+
  labs(x="group", y="relative abundance", title= "Lactobacillus Controles")+
  theme_pubclean()+scale_y_log10(breaks = c(0.0001,0.001,0.01,0.1,1))
ggsave("RelativeAbundancesLactobacilli_baselineControles.png")

barstaphbl<-ggplot(relabunStaphBL,aes(x=acne,y=rel_abundance))+
  geom_boxplot(alpha=0.5,fill='lightgrey',outlier.size = 0)+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=1,alpha=0.7,aes(col= subgroup, fill=subgroup))+
  scale_color_manual(values=c("#66a182","#2e4057","#00798c"))+
  scale_fill_manual(values=c("#66a182","#2e4057","#00798c"))+
  labs(x="group", y="relative abundance", title= "Staphylococcus-baseline")+
  theme_pubclean()+scale_y_log10(breaks = c(0.0001,0.001,0.01,0.1,1))+expand_limits(y=2)+theme(legend.position = "bottom")
ggsave("RelativeAbundancesStaphylococci_baseline.png")

barCutibl<-ggplot(relabunCutiBL,aes(x=acne,y=rel_abundance))+
  geom_boxplot(alpha=0.5,fill='lightgrey',outlier.size = 0)+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=1,alpha=0.7,aes(col= subgroup, fill=subgroup))+
  scale_color_manual(values=c("#66a182","#2e4057","#00798c"))+
  scale_fill_manual(values=c("#66a182","#2e4057","#00798c"))+
  labs(x="group", y="relative abundance", title= "Cutibacterium-baseline")+
  theme_pubclean()+scale_y_log10()+theme(legend.position = "bottom")
ggsave("RelativeAbundancesCutibacterium_baseline.png")
```
```{r,fig.width=10}
ggarrange(barstaphbl,BarStaphstudy,barCutibl,BarCutistudy,ncol=4,common.legend = T)
ggsave("barRelAbund.svg")
```


```{r}

write.csv2(get_abundances_extended(studym),"studym.csv")
write.csv2(get_abundances_extended(baselinem),"baselinem.csv")
write.csv2(get_abundances_extended(studyLacto),"studyLacto.csv")
write.csv2(baselineLacto$taxa,"baselineLacto.csv")
write.csv2(relabunLacCon,"relabunLacCon.csv")
write.csv2(relabunStaph,"relabunStaph.csv")
write.csv2(relabunCuti,"relabunCuti.csv")
write.csv2(relabunLac,"relabunLac.csv")
```
#Study

```{r}
get_bar_plot(studymg, geom_bar =F)+
  geom_bar(stat="identity", position="fill", aes(x=visit))+
  labs(x="Visit",title="Taxonomic composition per sample")+
  facet_wrap(~participant)+theme_minimal()
ggsave("TaxonomyStudy.png")

get_bar_plot(studymg, geom_bar =F)+
  geom_bar(stat="identity", position="fill", aes(x=visit))+
  labs(x="Visit",title="Taxonomic composition per sample")+
  facet_wrap(~participant)+theme_minimal()+theme(legend.position = "bottom",axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14), legend.text = element_text(size=11))
ggsave("TaxonomyStudy_2.png")

get_bar_plot(studynoLactog, geom_bar =F)+
  geom_bar(stat="identity", position="fill", aes(x=visit))+
  labs(x="Visit",title="Taxonomic composition per sample")+
  facet_wrap(~participant,nrow=1)+theme_minimal()
ggsave("TaxonomyStudy_LactoFiltered.png")



```

```{r,fig.width=10}
get_bar_plot(studymg, geom_bar =F)+
  geom_bar(stat="identity", position="fill", aes(x=visit))+
  labs(x="Visit",title="Taxonomic composition per sample")+
  facet_wrap(~participant,nrow=1)+theme_minimal()+theme(legend.position = "right",axis.text.x= element_text(size = 10, angle=90),axis.text.y= element_text(size = 12), axis.title=element_text(size=14), legend.text = element_text(size=11))
ggsave("BarGenusTax.svg")
```


#Alpha diversity

```{r}
baselinemg<-add_diversity_measures(baselinemg)
studymg<-add_diversity_measures(studymg)

ggplot(studymg$samples,aes(x=visit, y=div_observed))+
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.7)+ 
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y.. ),
                 width = .76, linetype = "dashed")+
  labs(x="Visit", y="Alpha diversity measure", title="Observed diversity") +
  stat_compare_means(comparisons = my_comparisons)+
  theme_pubclean()

ggplot(studymg$samples,aes(x=visit, y=div_inv_simpson))+
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.7)+ 
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y.. ),
                 width = .76, linetype = "dashed") +
  stat_compare_means(comparisons = my_comparisons)+
  labs(x="Visit", y="Alpha diversity measure", title="Inverse Simpson index")+
  theme_pubclean()

ggplot(baselinemg$samples,aes(x=acne, y=div_observed))+
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.7, aes(col=subgroup))+ 
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y.. ),
                 width = .76, linetype = "dashed") +
  stat_compare_means()+
  labs(x="Acne sample", y="Alpha diversity measure", title="Observed diversity")+
  theme_pubclean()

ggplot(baselinemg$samples,aes(x=acne, y=div_inv_simpson))+
  geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.7, aes(col=subgroup))+ 
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y.. ),
                 width = .76, linetype = "dashed") +
  stat_compare_means()+
  labs(x="Acne sample", y="Alpha diversity measure", title="Inverse Simpson index")+
  theme_pubclean()

```


Make figures for the relative abundances
```{r}
ggplot(relabunStaph, aes(x=comedones,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title =" Staphylococcus", x = "comedones", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)

ggplot(relabunStaph, aes(x=pandp,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title =" Staphylococcus", x = "inflammatory lesions", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)

ggplot(relabunLac, aes(x=comedones,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title ="Lactobacillus", x = "comedones", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)

ggplot(relabunLac, aes(x=pandp,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title =" Lactobacillus", x = "inflammatory lesions", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)

ggplot(relabunCuti, aes(x=comedones,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title =" Cutibacteria", x = "comedones", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)

ggplot(relabunCuti, aes(x=pandp,y=rel_abundance,colour=visit))+geom_point()+facet_wrap(~visit)+geom_smooth(method = "lm")+stat_cor(label.x.npc = "left",label.y = 0.7)+ labs(title =" Cutibacteria", x = "inflammatory lesions", y = "Relative abundance")+theme_pubclean()+expand_limits(y=-0.2)
```



```{r}
ggplot(studymg$samples,aes(x=visit, y=pandp))+geom_point()+facet_wrap(~participant)+theme_pubclean()
```

Make PCOA plot for the study and the baseline samples
```{r}
studymg<-add_pcoa(studymg)
ggplot(studymg$samples,aes(x=pcoa1,y=pcoa2,col=as.factor(visit)))+geom_point()+theme_minimal()+ labs(color='Visit') +theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14),legend.title=element_text(size=14))

baselinemg<-add_pcoa(baselinemg)
ggplot(baselinemg$samples,aes(x=pcoa1,y=pcoa2, col= acne))+geom_point()+theme_minimal()+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14),legend.title=element_text(size=14))
```

check relative abundances at baseline for delbrueckii group

```{r}

baselineLactoDb<-baselineLacto
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 1$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 2$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 3$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 6$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 7$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 9$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 18$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 21$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 23$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 25$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 26$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 28$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 29$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 30$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 32$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 33$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 35$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 38$"))
baselineLactoDb$taxa<-filter(baselineLactoDb$taxa, !str_detect(taxon_name, "Lactobacillus 39$"))
baselineLactoDb<-process_taxon_selection(baselineLactoDb)


baselineLactoDb<-add_rel_abundance(baselineLactoDb)
relabundDb<-get_abundances_extended(baselineLactoDb)

```

```{r}
ggplot(relabundDb,aes(x=acne,y=rel_abundance))+
  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=0.7)+
  stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.7)+
  stat_compare_means()+
  labs(x="group", y="relative abundance", title= "Delbrueckii")+
  theme_pubclean()+scale_y_log10(breaks = c(0.0001,0.001,0.01,0.1,1))+expand_limits(y=2)

```



```{r}
write.csv2(get_abundances_extended(baselinem),"baselinem.csv")
```

```{r}
#ggplot(delbrueckii,aes(x = group,y=RelAbun))+
#  geom_dotplot(binaxis = "y", stackdir = "center",dotsize=0.7)+
#  stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.7)+
#  stat_compare_means(label.y.npc= "bottom",label.x.npc= "middle")+
#  labs(x="group", y="relative abundance", title= "Delbrueckii group")+
#  theme_pubclean()+scale_y_log10()
```

Make altered figure for growth in figure 4
```{r}
growth_pac <- read.csv("~/Desktop/Miseq/analysis/V1V2run4/paper/growth_pac.csv", sep=";")
growth_pac$growth.1<-factor(growth_pac$growth.1,levels =c("No","Weak","Strong"))
```

```{r,fig.width=12,fig.height=1}
fig4dbottom<-ggplot(growth_pac,aes(x=visite,y=medium))+geom_tile(aes(fill=growth.1))+facet_wrap(vars(patient),nrow=1)+theme_minimal()+theme(text = element_text(size=18))+scale_fill_brewer(palette = "Blues")+theme(legend.position = "bottom")
fig4dbottom
```

```{r,fig.width=12}

ggarrange(fig4dtop,fig4dm,fig4dbottom,nrow=3)
ggsave("fig4d.svg")
```


```

